---
title: "p8105_hw3_xy2397"
author: "Xue Yang"
date: "10/6/2018"
output: github_document
---

```{r setup, include=FALSE}


library(tidyverse)
library(ggridges)
library(patchwork)
library(hexbin)
```

# Problem 1

**Load the data**

```{r}
library(p8105.datasets)
```

**Read and clean the data**

•	format the data to use appropriate variable names;
•	focus on the “Overall Health” topic
•	include only responses from “Excellent” to “Poor”
•	organize responses as a factor taking levels from “Excellent” to “Poor”

```{r}

brfss = 
  
  # read the data "brfss_smart2010" from the p8105.datasets package
  brfss_smart2010 %>% 
  
  # clean up variable names
  janitor::clean_names() %>% 
  
  # filter the data to only focus on "Overall Health" topic
  filter(., topic == "Overall Health") %>% 
  
   # mutate responses as a factor taking levels from “Excellent” to “Poor”
  mutate(., response = factor(response)) %>% 
  
  # include only responses from “Excellent” to “Poor”
  select(., year, locationabbr, locationdesc, response, data_value) %>%
  spread(key = "response", value = "data_value") %>% 
  
  # clean up variable names again
  janitor::clean_names()
  
```

**Answering the questions using this dataset**

•	In 2002, which states were observed at 7 locations?

```{r}
brfss_2002 = 
  brfss %>% 
  filter(year == "2002") %>%
  group_by(locationabbr) %>% 
  summarize(number = n()) %>% 
  filter(number == 7)


brfss_2002
```

So we can see that states "CT", "FL" and "NC" were observed at 7 locations.

•	Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

```{r}

  brfss %>% 
  group_by(year, locationabbr) %>% 
  summarize(number = n()) %>% 
  ggplot(aes(x = year , y = number, color = locationabbr)) +
  geom_line() +
  labs(
    title = "Spaghetti plot",
    x = "Year",
    y = "Number of observations in each state",
    caption = "Data from the p8105.datasets package"
  ) 
```


•	Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

```{r}
brfss %>% 
  filter(year %in% c("2002", "2006", "2010"), locationabbr == "NY") %>% 
  mutate(prop = excellent / (excellent + very_good + good + fair + poor)) %>% 
  group_by(year) %>% 
  summarize(mean = mean(prop, na.rm = TRUE),
            sd = sd(prop, na.rm = TRUE)) %>% 
  knitr::kable(digits = 3)
```

•	For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

```{r}
brfss %>% 
  mutate(exc_prop = excellent / (excellent + very_good + good + fair + poor)) %>%
  mutate(vgd_prop = very_good / (excellent + very_good + good + fair + poor)) %>%
  mutate(gd_prop = good / (excellent + very_good + good + fair + poor)) %>%
  mutate(fa_prop = fair / (excellent + very_good + good + fair + poor)) %>%
  mutate(pr_prop = poor / (excellent + very_good + good + fair + poor)) %>% 
  group_by(year, locationabbr) %>% 
  summarize(exc_prop_average = mean(exc_prop,na.rm = TRUE),
            vgd_prop_average = mean(vgd_prop, na.rm = TRUE),
            good_prop_average = mean(gd_prop, na.rm = TRUE),
            fair_prop_average = mean(fa_prop, na.rm = TRUE),
            poor_prop_average = mean(pr_prop, na.rm = TRUE)) %>% 
  gather(., key = "response", value = "average_prop", exc_prop_average:poor_prop_average) %>%    ggplot(aes(x = year, y = average_prop, color = locationabbr)) +
  geom_point()+
  facet_grid(~response) +
  labs(
    title = "Five-panel plot",
    x = "Year",
    y = "Proportion of state-level averages",
    caption = "Data from the p8105.datasets package"
  ) +
  viridis::scale_color_viridis(
    name = "State",
    discrete = TRUE
    ) + 
  theme(legend.position = "bottom")
```



# Problem 2

**Load and read the data**

```{r}
data("instacart")
```


The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. 

The dataset "Instacart" contains online grocery orders from instacart users.The size of the dataset is `r nrow(instacart)` rows x `r (ncol(instacart))` columns. 

Then, do or answer the following (commenting on the results of each):
•	How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(number = n()) %>% 
  mutate(aisle_rank = min_rank(desc(number))) %>% 
  filter(aisle_rank == "1")
```
So there are totally 134 aisles there, and fresh vegetables is the most items ordered from.

•	Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarise(number = n()) %>% 
  ggplot(aes(x = aisle, y = number)) +
  geom_point()
```


•	Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits" ) ) %>% 
  group_by(aisle) %>% 
  summarize(number = n()) %>% 
  knitr::kable(digits = 3)

```


•	Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).


```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour_day = mean(order_hour_of_day, na.rm = TRUE)) %>% 
  spread(key = "order_dow", value = "mean_hour_day") %>% 
  knitr::kable(digits = 3)
  
```

# Problem 3
```{r}
data("ny_noaa")
ny_noaa
```

The goal is to do some exploration of this dataset. 

To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. 

This dataset "ny_noaa" provides some weather data, it contains records from 747 different weather stations for about datas from 10,957 obervational  days. The size of the dataset is `r nrow(ny_noaa)` rows x `r (ncol(ny_noaa))` columns. The key variables are id (weather station ID), date (date of observation), prcp (precipitation for tenths of mm), snow (snowfall for mm), snwd (snow depth for mm) and tmax and tmin (maximum and mimimum temperature for tenths of degrees C). At last, we calculate the proportion of missing value fro each vairable, we have  0.0561958 for prop, 0.146896 for snow, 0.2280331 for snwd, 0.4371025 for tmax and 0.4371264 for tmin.

```{r}
ny_noaa %>% 
  group_by(id) %>% 
  summarise(number = n())

ny_noaa %>% 
  group_by(date) %>% 
  summarise(number = n()) 

prcp_na =
  ny_noaa %>% 
  filter(is.na(prcp)) %>% 
  nrow()
prcp_prop = prcp_na/nrow(ny_noaa)
prcp_prop

snow_na =
  ny_noaa %>% 
  filter(is.na(snow)) %>% 
  nrow()
snow_prop = snow_na/nrow(ny_noaa)
snow_prop

snwd_na =
  ny_noaa %>% 
  filter(is.na(snwd)) %>% 
  nrow()
snwd_prop = snwd_na/nrow(ny_noaa)
snwd_prop

tmax_na =
  ny_noaa %>% 
  filter(is.na(tmax)) %>% 
  nrow()
tmax_prop = tmax_na/nrow(ny_noaa)
tmax_prop

tmin_na =
  ny_noaa %>% 
  filter(is.na(tmin)) %>% 
  nrow()
tmin_prop = tmin_na/nrow(ny_noaa)
tmin_prop
```


Then, do or answer the following (commenting on the results of each):

•	Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?

```{r}
noaa = 
  ny_noaa %>%
  separate(date, into = c("year", "month", "day"), sep = "-")
```

```{r}
noaa %>%
  count(snow) %>% 
  top_n(5) 

```
So for snowfall, 0 is the most commonly observed value, since 0 has been observed for 2008508 times from the dataset. The second commonly observed value id 25 and it was observed for 31022 times.

•	Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r}
noaa %>% 
  filter(month %in% c("01", "07")) %>% 
  mutate(tmax = as.numeric(tmax)) %>% 
  group_by(id, month) %>% 
  summarize(average_tmax = mean(tmax, na.rm = TRUE)) %>% 
  filter(!is.na(average_tmax)) %>% 
  ggplot(aes(x = id, y = average_tmax, color = month)) +
  geom_point(alpha = 0.5) +
  geom_line() +
  facet_grid(~month) +
  labs(
    title = "Average max temperature two-panel plot",
    x = "Weather Station",
    y = "Average max temperature across year (tenths of degrees CC)",
    caption = "Data from the p8105.datasets package"
  ) +
  theme(legend.position = "bottom")
  
```
When draw a plot for average max temperature for in January and in July in each station across years, we first delete all the missing value. From the two-panel plot we can easily find that the average temperature recorded from each station in January is lower than in July, it's obviously, since it's winter in January and summer in July. And there are also some outliers in January which may be explained that there are few countries and territories are especially cold in January.

•	Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
tmax_tmin =
  noaa %>% 
  filter(!is.na(tmax), !is.na(tmin)) %>% 
  ggplot(aes(x = tmax, y = tmin)) +
  geom_hex(bins = 300) +
  labs(
    title = "Tmax vs Tmin",
    x = "Maximum temperature",
    y = "Minimum temperature",
    caption = "Data from the p8105.datasets package"
  ) +
  scale_x_discrete(
    limits = c(-250, 250)
  ) +
  theme(legend.position = "bottom")

snowfall =
  noaa %>% 
  filter(snow %in% 1:99) %>% 
  ggplot(aes(x = snow, fill = year)) +
  geom_histogram(alpha = 0.5) +
  labs(
    title = "Distribution of snowfall",
    x = "Snowfall values (mm)",
    y = "Frequency",
    caption = "Data from the p8105.datasets package"
  ) +
  theme(legend.position = "bottom")

tmax_tmin + snowfall
```

So when we deal with the 