---
title: "p8105_hw3_xy2397"
author: "Xue Yang"
date: "10/6/2018"
output: github_document
---

```{r setup, include=FALSE}


library(tidyverse)
library(ggridges)
```

# Problem 1

**Load the data**

```{r}
library(p8105.datasets)
```

**Read and clean the data**

•	format the data to use appropriate variable names;
•	focus on the “Overall Health” topic
•	include only responses from “Excellent” to “Poor”
•	organize responses as a factor taking levels from “Excellent” to “Poor”

```{r}

brfss = 
  
  # read the data "brfss_smart2010" from the p8105.datasets package
  brfss_smart2010 %>% 
  
  # clean up variable names
  janitor::clean_names() %>% 
  
  # filter the data to only focus on "Overall Health" topic
  filter(., topic == "Overall Health") %>% 
  
   # mutate responses as a factor taking levels from “Excellent” to “Poor”
  mutate(., response = factor(response)) %>% 
  
  # include only responses from “Excellent” to “Poor”
  select(., year, locationabbr, locationdesc, response, data_value) %>%
  spread(key = "response", value = "data_value") %>% 
  
  # clean up variable names again
  janitor::clean_names()
  
```

**Answering the questions using this dataset**

•	In 2002, which states were observed at 7 locations?

```{r}
brfss_2002 = 
  brfss %>% 
  filter(year == "2002") %>%
  group_by(locationabbr) %>% 
  summarize(number = n()) %>% 
  filter(number == 7)


brfss_2002
```

So we can see that states "CT", "FL" and "NC" were observed at 7 locations.

•	Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

```{r}

  brfss %>% 
  group_by(year, locationabbr) %>% 
  summarize(number = n()) %>% 
  ggplot(aes(x = year , y = number, color = locationabbr)) +
  geom_line() +
  labs(
    title = "Spaghetti plot",
    x = "Year",
    y = "Number of observations in each state",
    caption = "Data from the p8105.datasets package"
  ) 
```


•	Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

```{r}
brfss %>% 
  filter(year %in% c("2002", "2006", "2010"), locationabbr == "NY") %>% 
  mutate(prop = excellent / (excellent + very_good + good + fair + poor)) %>% 
  group_by(year) %>% 
  summarize(mean = mean(prop, na.rm = TRUE),
            sd = sd(prop, na.rm = TRUE)) %>% 
  knitr::kable(digits = 3)
```

•	For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

```{r}
brfss %>% 
  mutate(exc_prop = excellent / (excellent + very_good + good + fair + poor)) %>%
  mutate(vgd_prop = very_good / (excellent + very_good + good + fair + poor)) %>%
  mutate(gd_prop = good / (excellent + very_good + good + fair + poor)) %>%
  mutate(fa_prop = fair / (excellent + very_good + good + fair + poor)) %>%
  mutate(pr_prop = poor / (excellent + very_good + good + fair + poor)) %>% 
  group_by(year, locationabbr) %>% 
  summarize(exc_prop_average = mean(exc_prop,na.rm = TRUE),
            vgd_prop_average = mean(vgd_prop, na.rm = TRUE),
            good_prop_average = mean(gd_prop, na.rm = TRUE),
            fair_prop_average = mean(fa_prop, na.rm = TRUE),
            poor_prop_average = mean(pr_prop, na.rm = TRUE)) %>% 
  gather(., key = "response", value = "average_prop", exc_prop_average:poor_prop_average) %>%    ggplot(aes(x = year, y = average_prop, color = locationabbr)) +
  geom_point()+
  facet_grid(~response) +
  labs(
    title = "Five-panel plot",
    x = "Year",
    y = "Proportion of state-level averages",
    caption = "Data from the p8105.datasets package"
  ) +
  viridis::scale_color_viridis(
    name = "State",
    discrete = TRUE
    ) + 
  theme(legend.position = "bottom")
```



# Problem 2

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):
•	How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(number = n()) %>% 
  mutate(aisle_rank = min_rank(desc(number))) %>% 
  filter(aisle_rank == "1")
```
So there are totally 134 aisles there, and fresh vegetables is the most items ordered from.

•	Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarise(number = n()) %>% 
  ggplot(aes(x = aisle, y = number)) +
  geom_point()
```


•	Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits" ) ) %>% 
  group_by(aisle) %>% 
  summarize(number = n()) %>% 
  knitr::kable(digits = 3)

```


•	Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).


```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour_day = mean(order_hour_of_day, na.rm = TRUE)) %>% 
  spread(key = "order_dow", value = "mean_hour_day") %>% 
  knitr::kable(digits = 3)
  
```


